---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
### 1. Load the data (unzip the original zip file and get activity.csv)
```{r GetData,echo=TRUE}
data<-read.csv("activity.csv")
#Set Locale thus date column could be judged by weekday's name
Sys.setlocale('LC_ALL','English')
```
### 2. Process/transform the data into a format suitable for analysis.

```{r Transform,echo=TRUE}

# convert column "interval" into factor:
data <- transform(data, interval = factor(interval)) 

```
## What is mean total number of steps taken per day
```{r GetMeanStepsPerDay1,fig.height=8,fig.width=14, echo=TRUE}
#Remove warning message when plotting with missing value
options(warn=-1)
#install.packages("ggplot2")
library(ggplot2)
StepSumPerDay <- tapply(data$steps, data$date, FUN = sum)
newDF<-data.frame(date=names(StepSumPerDay), stepSum=StepSumPerDay)
ggplot(newDF,aes(x=date,y=stepSum))+
        geom_histogram(stat='identity', fill="lightblue", colour="black")+
        theme(axis.text.x = element_text(angle=60, hjust=1))
#if base graphic needed, then use code below:
# barplot(StepSumPerDay)

summary(StepSumPerDay,rm.na=T)
```
## What is the average daily activity pattern?
```{r GetStepsPerInterval,fig.height=8,fig.width=14, echo=TRUE}
#install.packages("plyr")
library(plyr)
StepsPerInterveal<-ddply(data, c("interval"),summarize, StepMean=mean(steps,na.rm = TRUE))

with(StepsPerInterveal, plot(interval, StepMean,type="l",xlab = "Interval",ylab="Averaged Steps",col="blue")) 

lines(StepsPerInterveal$interval, StepsPerInterveal$StepMean,type="l",xlab = "Interval",ylab="Averaged Steps",col="blue")

```

### Base on the plot, we can apprently observe that the interval 845 (it's 8:45AM) has the maximum averaged number of steps each day.

### Imputing missing values
### 1.Calculate and report the total number of missing values in the dataset 
###   ==develop a function NAreport to report NA distribution 
###   ==NAtype=c("NA","","N/A","NULL","--","NONE")
```{r MissingDataImputation,fig.height=8,fig.width=14, echo=TRUE}
NAsum<-sum(is.na(data$steps))
```
The total number of missing values is: `r NAsum`

### 2.Devise a strategy for filling in all of the missing values in the dataset. 
### The strategy does not need to be sophisticated. For example, you could use the
### mean/median for that day, or the mean for that 5-minute interval, etc.
### Here, we use the mean for that 5-minute interval to impute NAs.

### 3.Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r NewDataWithNAfilled,echo=TRUE}
dataNew<-data
dataNew$steps[is.na(dataNew$steps)]<- StepsPerInterveal$StepMean[match(dataNew$interval, StepsPerInterveal$interval)]
```

### 4.Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day
```{r GetMeanStepsPerDayfromNewData,fig.height=8,fig.width=14, echo=TRUE}
StepSumPerDayNew <- tapply(dataNew$steps, dataNew$date, FUN = sum)
newDF<-data.frame(date=names(StepSumPerDayNew), stepSum=StepSumPerDayNew)
ggplot(newDF,aes(x=date,y=stepSum))+
        geom_histogram(stat='identity', fill="lightblue", colour="black")+
        theme(axis.text.x = element_text(angle=60, hjust=1))
summary(StepSumPerDayNew,rm.na=T)
```

###With Filled missing value, we could find that the new Mean value of steps tasken per day is same as the original one while the Median is a little different (10770 vs 10760) but could be ignored. So the only impact of the filled missing data is on the special day in which there are missing values and get an obvious increasing value in the plot. Anyway, it doesn't impact the data in whole.

## Are there differences in activity patterns between weekdays and weekends?

```{r AddWeekendColumn,fig.height=12,fig.width=16, echo=TRUE}
# convert column "date" into datetime format:
library(lattice)
dataNew$date<-strptime(dataNew$date,"%Y-%m-%d") 
#1.Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
dataNew$day = ifelse(weekdays(dataNew$date) == "Saturday" | weekdays(dataNew$date) == "Sunday", "weekend", "weekday")
#Convert the new variable into factor class
dataNew <- transform(dataNew, day = factor(day))

#2.Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). See the README file in the GitHub repository to see an example of what this plot should look like using simulated data.

#Remove date variable in order to run ddplyr function
#dataNew<-dataNew[ , !names(dataNew) %in% c("date")] 

#get the average datasets both for weekday and weekend
StepsPerIntervealDay<-ddply(dataNew, c("interval","day"),summarize, StepMean=mean(steps,na.rm = TRUE))
#StepsPerIntervealWeekday<-StepsPerIntervealDay[StepsPerIntervealDay$day=="weekday",]
#StepsPerIntervealWeekend<-StepsPerIntervealDay[StepsPerIntervealDay$day=="weekend",]

xyplot(StepMean ~ interval | day , data = StepsPerIntervealDay, type="l",main="Averaged Steps on Weekday and Weekend ",ylab = "Averaged Steps", xlab = "Interval", scales=list(y=list(tick.number=15),x=list(tick.number=25,at = seq(1, 288, 15))), layout = c(1, 2))

```